---
title: "KandruB.CS5200.SQL-Queries.Sp24.Rmd"
output:
  html_document:
    df_print: paged
---
```{r connect2DB, echo=F, warning=F}
library(RSQLite)
library(DBI)
dbcon <- dbConnect(RSQLite::SQLite(),dbname = "MediaDB.db")
knitr::opts_chunk$set(connection = "dbcon")
```

```{sql connection=dbcon, echo=F}
PRAGMA foreign_keys = ON
```
```{r}
dbListTables(dbcon)
dbListFields(dbcon,'genres')
```
1) What are the last name, first name, title, and hire date of all employees, sorted by last name?
```{sql connection=dbcon}
SELECT LastName,FirstName,Title,HireDate
FROM employees
ORDER BY LastName;
```
2) What are the names of each genre and the number of distinct albums in that genre? 
```{sql connection=dbcon}
SELECT g.Name, COUNT(DISTINCT a.AlbumId) AS NumberOfAlbums
FROM genres g
JOIN tracks t ON g.GenreId = t.GenreId
JOIN albums a ON t.AlbumId = a.AlbumId
GROUP BY g.Name;
```
3) What are the distinct media types?
```{sql connection=dbcon}
SELECT DISTINCT Name 
FROM media_types;
```
4) List the number of tracks by genre, ordered from most to least, restricted to those genres for which there are at least four tracks.
```{sql connection=dbcon}
SELECT g.Name, COUNT(t.TrackID) AS NumberOfTracks
FROM genres g
JOIN Tracks t ON g.GenreID=t.GenreID
GROUP BY g.Name 
HAVING COUNT(t.TrackID)>=4
ORDER BY NumberOfTracks DESC;
```
5) What is the last name and first name of the youngest employee?
Hint: SQLite does not support DATEDIFF but does support STRFTIME (see tutorial Links to an external site.). You need to only take year and month into account.
```{sql connection=dbcon}
SELECT LastName,FirstName 
FROM employees
ORDER BY STRFTIME('%Y-%M',BirthDate) DESC
LIMIT 1;
```
6) What are the total number of unique purchases for each Brazilian state (e.g., MG, SP, etc.), ordered alphabetically by state, and in which there were at least two purchases?
```{sql connection=dbcon}
SELECT c.state, COUNT(DISTINCT i.InvoiceID) AS NumberOfPurchase
FROM customers c
JOIN invoices i ON c.CustomerID=i.CustomerID
WHERE c.Country = "Brazil"
GROUP BY c.state 
HAVING COUNT(DISTINCT i.InvoiceID)>=2
ORDER BY c.State;
```
7) How many employees never worked with a single customer, i.e., they appear in the employees table but not in the customers table?
```{sql connection=dbcon}
SELECT count(EmployeeId)
FROM employees
WHERE EmployeeId NOT IN (SELECT SupportRepId FROM customers);
```
8) How many albums contain the substring "Classic" but not "Hard" in the album title?
```{sql connection=dbcon}
SELECT COUNT(*)
FROM albums
WHERE Title LIKE "%Classic%" AND Title NOT LIKE "%Hard%";
```
9) What is the total number of distinct genres?
```{sql connection=dbcon}
SELECT DISTINCT COUNT(*)
FROM genres;
```
10) What is the total "time" of all tracks per album type in hours, i.e., add up the "Milliseconds" column in tracks? Display the media type and the time in hours rounded to one decimals. Only consider media types with at least one track.
```{sql connection=dbcon}
SELECT m.Name, ROUND(SUM(t.Milliseconds)/3600000.0, 1) AS TimeInHours
FROM tracks t
JOIN media_types m ON t.MediaTypeId = m.MediaTypeId
GROUP BY m.MediaTypeId
HAVING COUNT(t.trackId)>=1 
ORDER BY TimeInHours;
```
11) What is the title and artist name of the longest album, in terms of minutes of playtime of all tracks on that albums?
```{sql connection=dbcon}
SELECT alb.Title, art.Name, t.Milliseconds
FROM albums alb
JOIN artists art ON alb.ArtistId = art.ArtistId
JOIN tracks t ON alb.AlbumId = t.AlbumId
GROUP BY alb.AlbumId
ORDER BY SUM(t.Milliseconds) DESC
LIMIT 1;
```
12) Which genre has the least number of tracks?
```{sql connection=dbcon}
SELECT g.Name AS GenreName, t.Composer
FROM genres g
JOIN tracks t ON g.GenreID=t.GenreId
GROUP BY g.GenreID
ORDER BY COUNT(t.TrackID) 
LIMIT 1;
```
13) Which media format has the fewest tracks?
```{sql connection=dbcon}
SELECT m.Name AS MediaFormat
FROM media_types m
JOIN tracks t ON m.MediaTypeID=t.MediaTypeId
GROUP BY m.MediaTypeId
ORDER BY COUNT(t.TrackID) 
LIMIT 1;
```
14) For each country, what is the total number of unique customers and the total spent?
```{sql connection=dbcon}
SELECT c.Country, COUNT(DISTINCT c.CustomerId) AS NumberOfCustomers,
SUM(i.Total) AS TotalSpent 
FROM customers c
JOIN invoices i ON c.CustomerId = i.CustomerId
GROUP BY c.Country;
```
15) What is the total amount spent by all customers who do not live in South America?
```{sql connection=dbcon}
SELECT SUM(ii.UnitPrice * ii.Quantity) AS TotalSpent
FROM invoices AS i
JOIN customers AS c ON i.CustomerId = c.CustomerId
JOIN invoice_items AS ii ON i.InvoiceId = ii.InvoiceId
WHERE c.Country NOT IN ('Brazil', 'Argentina', 'Chile', 'Colombia', 'Peru', 'Venezuela', 'Ecuador', 'Bolivia', 'Paraguay', 'Uruguay', 'Guyana', 'Suriname');
```
16) What is the difference in spending on any type of MPEG vs any type of AAC track?
```{sql connection=dbcon}
SELECT(
  (SELECT SUM(ii.UnitPrice * ii.Quantity)
   FROM invoice_items ii
   JOIN tracks t ON ii.TrackId = t.TrackId
   JOIN media_types mt ON t.MediaTypeId = mt.MediaTypeId
   WHERE mt.Name LIKE '%MPEG%') -
  (SELECT SUM(ii.UnitPrice * ii.Quantity)
   FROM invoice_items ii
   JOIN tracks t ON ii.TrackId = t.TrackId
   JOIN media_types mt ON t.MediaTypeId = mt.MediaTypeId
   WHERE mt.Name LIKE '%AAC%')) AS Difference;
```
17) Which country has the most customers?
```{sql connection=dbcon}
SELECT Country, COUNT(*) AS MostCustomers
FROM customers
GROUP BY Country
ORDER BY MostCustomers DESC
limit 1;
```
18) How many tracks did "Amy Winehouse" record?
```{sql connection=dbcon}
SELECT art.Name, COUNT(alb.AlbumId) AS NumberOfTracks
FROM artists art
JOIN albums alb ON art.ArtistId=alb.ArtistId
JOIN tracks t ON alb.AlbumId = t.AlbumId
WHERE art.Name = "Amy Winehouse";
```
19) What is the name of the artist who earned the most, i.e., whose aggregate total of all tracks sold is the highest?
```{sql connection=dbcon}
SELECT a.Name, COUNT(i.UnitPrice*i.Quantity) AS Earnings
FROM invoice_items i
JOIN tracks t ON i.TrackId = t.TrackId
JOIN albums alb ON t.AlbumId = alb.AlbumId 
JOIN artists a ON alb.ArtistId = a.ArtistId
GROUP BY a.ArtistId
ORDER BY Earnings DESC
LIMIT 1;
```
20) Which artist is the most popular in Paris? Popularity is defined in this context as the artist who was purchased most often by someone living in Paris.
```{sql connection=dbcon}
SELECT art.Name, COUNT(*) AS MostPurchases
FROM invoices i 
JOIN customers c ON i.CustomerId = c.CustomerId
JOIN invoice_items ii ON i.InvoiceId = ii.InvoiceId
JOIN tracks t ON ii.TrackId = t.TrackId
JOIN albums a ON t.AlbumId = a.AlbumId
JOIN artists art ON a.ArtistId = art.ArtistId
WHERE c.city = "Paris"
GROUP BY art.ArtistId
ORDER BY MostPurchases DESC
LIMIT 1;
```